---
title: "Tibbletime"
author: "Jamie Afflerbach"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
# Install from either CRAN or GH
# install.packages("tibbletime")
devtools::install_github("business-science/tibbletime")

install.packages("lubridate")

# Load libraries
library(tidyverse)
library(tibbletime)
library(lubridate)
```

## Data

```{r}

data <- read_csv("data/sharks.csv")

View(data)
class(data)
```

First we need to turn this `tbl_df` into a `tbl_time`. You use the `as_tbl_time()` function for this.

```{r}
sharks_tbl <- as_tbl_time(sharks, index = Date)
```

`tibbletime` doesn't recognize the date column. If you look at `sharks` you see that under "Date" it says `chr` meaning R sees it as a character, not a date. Using the `lubridate` package we can change these dates into the correct format. Specifically, the `mdy()` function since our dates are currently in that format.

```{r}

sharks <- data %>%
  mutate(Date = mdy(Date))

sharks
```

Now it shows this column as a date.

Let's try that again.

```{r}
sharks_tbl <- as_tbl_time(sharks, index = Date) %>%
  arrange(Date) %>%
  mutate(Depth = Depth * -1) #changing depth to be positive

class(sharks_tbl)
```

Now we can parse this dataset for what we want.

Let's look at mean depth and temperature in summer vs winter months.

You can treat these tbl_time dataframes just like regular dataframes.

```{r}
range(sharks_tbl$Date)
```

Date specific
```{r}
sharks_tbl %>%
  filter_time("2012-06-21" ~ "2012-09-21")
```

More broad (looking for months not specific dates)
```{r}
sharks_tbl %>%
  filter_time("2012-06" ~ "2012-09")
```

Start of the time series to specified date/month/year
```{r}
sharks_tbl %>%
  filter_time("start" ~ "2012-12") #everything before January 2013.
```

## Analysis with time_tbl

Mean depth and temp of shark dives in summer of 2012.

```{r}

sharks_tbl %>%
  filter_time("2012-06" ~ "2012-09") %>%
  group_by(ID) %>%
  summarize(mean_depth = mean(Depth),
            mean_temp  = mean(Temp))

```

Grouping with tibbletime

```{r}
collapse_by(sharks_tbl, period = "monthly") %>%
  group_by(Date) %>%
  summarise(mean_depth = mean(Depth),
            mean_temp  = mean(Temp))
```

```{r}
collapse_by(sharks_tbl, period = "monthly") %>%
  group_by(Date) %>%
  summarise(mean_depth = mean(Depth),
            mean_temp  = mean(Temp)) %>%
  ggplot() +
  geom_line(aes(x = Date, y = mean_temp)) +
  ylab("Mean dive depth (m)")
```


